Modeling thermal dust emission with two components: application to the {\it PLANCK} HFI maps

### 2014-08-10 ms.tex
removed all comments

% might consider changing the word `specifies' in penultimate sentence

% specify early on that this is 'Galactic dust', 'dust' is kind of a vague
% term ... on the other hand the Planck dust paper is pretty lax about this

% mention more explicitly the enhanced temperature resolution relative to SFD??

@@

% mention re-radiation somewhere ?

@@

% DIRBE beam solid angles:
% http://lambda.gsfc.nasa.gov/product/cobe/dirbe_products.cfm

@@

% additional sentence suggested by doug : 
% The fact that the range of optical
% properties of small silicate particles measured in the lab (Agladze et
% al. 1996) spans the range needed by the FDS99 model makes plausible
% that the distinct species are real.

@@

\section{Data}
\label{sec:data}
All \PLANCK~data products utilized throughout this work are drawn from the 
\PLANCK~2013 release \citep{planck2013}. Specifically, we have made use 
of all six of the zodiacal light corrected HFI intensity maps
\citep[\texttt{R1.10\_nominal\_ZodiCorrected},][]{planckzodi}. Our 
full-resolution (6.1$'$ FWHM) SED fits neglect the two lowest HFI frequencies, 
100 and 143 GHz, as these have FWHM of 9.66$'$ and 7.27$'$ respectively.
% in last sentence, reference Table 1 ?

@@

The following subsections detail the processing steps we have applied to 
isolate Galactic dust emission in the \PLANCK~maps in preparation for SED
fitting. % reference fitting section at end of this sentence ?

@@

% NEED TO MENTION SMOOTHING OF THE SMICA MAP TO 6.1 ARMCIN !!!!!!

\subsection{Compact Sources}
\label{sec:ptsrc}
% this compact source section still needs a lot of work
After subtracting the SMICA CMB model, we interpolate over compact sources, 
including both point sources and resolved galaxies. Removing compact sources at
this stage is important as it prevents contamination of compact-source-free 
pixels in our downstream analyses which require smoothing of the \PLANCK~maps. 
SFD carefully removed point sources and galaxies from the \verb|i100| map 
everywhere outside of $|b|$$<$$5^{\circ}$. We do not perform any further 
modifications of the \verb|i100| map to account for compact sources. To mask 
compact sources in the \PLANCK~217-857 GHz maps, we use the SFD compact source 
mask. At 100, 143 GHz we use the compact source masks provided by the 
\PLANCK~collaboration in the file \verb|HFI_Mask_PointSrc_2048_R1.10.fits|. 
Given our pixelization (see $\S$\ref{sec:pix}), 1.56\% of pixels are masked 
at 217-857 GHz (1.05\%, 1.02\% at 100, 143 GHz)

% best to refer to an appendix for details of the point source masking ?

@@

% might consider quoting rounder values in text, most precise values are:
% 100 GHz, J=1->0 : 1.478e-5, 217 GHz, J=2->1 : 4.585e-5, 
% 353 GHz, J=3->2 : 1.751e-4

@@

% maybe a final sentence or two warning that, to whatever extent
% CO line ratios vary over the sky and/or other molecules make a substantial
% contribution to the observed Planck intensities, our molecular emission
% correction will be insufficient

% for CO line list see ~/wise/pro/planck_co_lines.pro

@@

% for quantitative assessment of the zero-point difference due to
% flattening of 857 GHz vs. HI, see 
% ~/Desktop/dust/planck_hi/piecewise.pdf (on my Mac)

% -> at low HI, slope decreases by a factor of 1.88
%    the resulting difference in 857 GHz zero points is 0.373 MJy/sr

@@

% maybe state approximately what 0.37 MJy/sr at 857 GHz equates to in EBV
% --> did a quick calculation and it's only ~0.0118 mags ... thought it
%     would be more ??

@@

% i evaluated the `formal statistical error' using the covar keyword in
% hogg_iter_linfit

@@

% what HI intensity does the 2.15 MJy/sr 857 GHz correspond to ?
%  ---> pretty high actually

@@

% mention how offset/dipole values compare to those derived in the Planck
% team paper ?

@@

% maybe a sentence to transition from preprocessing -> modeling
% e.g. With pre-processing completed, we turned to modeling the 
% thermal dust emission.

@@

% kind of feel like i'm re-writing the introduction here ??

% could add a last sentence 'ultimately this allows us to make high-resolution
% foreground and extinction predictions' and/or even find a way to mention
% other predictions, such as polarization/R_v

% in the above paragraph, should I mention the Planck collaboration result
% of beta_mm vs. beta_fir being different, with (beta_fir-beta_mm) > 0

% should i also try to justify why it's worth doing this fit even
% if it's possible to just claim different beta's at different frequencies?

@@

% With $f_{1}$ = 0.0363, $\beta_1$ = 1.67, $\beta_2$ = 2.70, and $q_1/q_2$ = 
% 13.0.

% re: following paragraph --- in the code nu_0 is actually 2997.92458
% because this is the frequency with respect to which DIRBE color correction
% K_100 (and my color correction as well) is referenced

% not sure if `Planck function' is the correct jargon

@@

% Importantly, the emissivity power law indices $\beta$ of the two species 
% differ.

@@

%%%%

@@

% would be great to have uncertainties on each parameter as well (in the
% Planck case at least...) !!!!!!!!!!!!!!!

%%%%

@@

%explain that we take the ``hot'' dust temperature to be T2 rather than T1

% maybe point to the SED fit summary figure as an example of what
%

@@

% is delta(nu)/nu conventionally (nu_off-nu_on)/nu_c or 1/2 of that ?

@@

% integral limits ??? check planck paper to see what they did
% -> planck spectral response paper just leaves the integrals indefinite

@@

%Again, since DIRBE-quoted fluxes also follow the 
%$\alpha$=$-1$ convention, Equation \ref{equ:bpcorr} remains applicable.

% what is the reference for the dirbe

%The color correction factors
%of Equation \ref{equ:bpcorr} are applied in the sense that 
%$b_{\nu_i}(T ,\beta)\times M_{\nu_{i,c}}$ is a prediction suitable for 
%direct comparison to the \PLANCK~data products.

@@

%Note that, as in FDS99, $f_2$=(1$-$$f_1$) by convention. 
The color correction of Equation \ref{equ:bpcorr} therefore allows us to 
predict $\tilde{I}_{\nu_i}$ by computing monochromatic flux densities at the 
central frequency $\nu_{i,c}$ and then multiplying by factors 
$b_{\nu_i}(T, \beta)$. In practice, we interpolated the color corrections off 
of a set of precomputed, one-dimensional lookup tables each listing 
$b_{\nu_i}(T, \beta)$ for a single $\beta$ value as a function of $T$. We thus
avoided the need to interpolate in both $\beta$ and $T$ by computing only a 
small set of one dimensional correction factors for the particular set of 
$\beta$ values of interest (e.g. $\beta$=1.67, 2.7, 1.63, 2.82 ..., see Table 
\ref{tab:global}). This color correction approach makes the MCMC sampling 
described in $\S$\ref{sec:mcmc} much more computationally efficient by 
circumventing the need to perform the integral in the numerator of Equation 
\ref{equ:bpcorr} on-the-fly for each proposed dust temperature. We have chosen 
to compute the color corrections on a per-MBB basis because this approach is 
very versatile; all possible two-component (and single-MBB) models are linear 
combinations of MBBs, so we can apply all of our color correction machinery 
even when we allow parameters other than temperature (e.g. $f_1$) to vary and 
thereby modify the dust spectrum shape.

% need some sort of clarification about the fact that, in the parts of the 
% analysis where beta floats, i actually do interpolate color corrections 
% off of a 2d grid in (T, beta)

% We note that it is not clear which bandpass should properly be ascribed to 
% the \IRAS/DIRBE composite 100$\mu$m maps under consideration.

@@

%is the old sentence below worth reinstating ?

%To reiterate, for 
%$\nu_i$$\le$353 GHz, $u_{\nu_i}(T_k, \beta_k)$ has units of 
%(MJy/sr)/$K_{CMB}$,and therefore $\tilde{I}_{\nu_{i}}$ will have different 
%units depending on the band $\nu_i$, though $\tilde{I}_{\nu_{i}}$ always has 
%the same units as the native band $\nu_i$ map.

% check that RIMO file i've used is actually the most up-to-date version

% extremely abrupt transition here...
\section{Global Model Parameters}
\label{sec:global}

@@

At 1250, 2141 and 3000 GHz, we use the SFD-reprocessed DIRBE maps. For each
DIRBE band, we determine $s_{i, 857}$ by performing a linear fit to DIRBE as a 
function of \PLANCK~857 GHz, after both have been zodiacal light subtracted and
smoothed to $1^{\circ}$ FWHM, also restricting to the low HI mask of 
$\S$\ref{sec:relzero} (see Figure \ref{fig:dirbe_slopes}).
% re-mention nside=64 or just bag it?
% restricted to same low HI mask as in Sect. 3.5.2 analysis
% sentence pointing to figure showing the regressions

@@

% maybe a table listing the correlation slopes and their uncertainties
% would be useful here

@@

% add error bars to this example SED plot

@@

% overplot ellipses showing 1 sigma errors in posterior grayscales

@@

% sentence or two mentioning reduced resolution fits where we sample in
% three dimensions (f_1, T_2, tilde{I}_{545})

@@

% be careful about notational consistency with e.g. tildes, etc.. in this sect

@@

%\begin{equation} \label{equ:like}
%\mathcal{L}(\vec{I}~|\tilde{I}_{545}, T_2) = \displaystyle\prod\limits_{i}\mathcal{N}(\tilde{I}_{\nu_{i}}|I_{\nu_{i}}, \sigma_{\nu_i})
%\end{equation}

@@

%be more careful about notation re: color corrections

@@

% is I_nu on RHS inclusive of offsets or is it the ``pure'' thermal dust
% intensity

% need to make sure to cite planckcalib paper

@@

% mention that each of the error terms is always rescaled properly according
% to smoothing/rebinning and relevant power spectrum

@@

%Some of 
%the most striking features occur near the north Galactic pole, and to 
%highlight these regions of unusually high $f_1$, we display the northern
%Galactic hemisphere in Lambert projection in Figure \ref{fig:f1}.

@@

% mention that when t floats, we find a lower T by ~ 1 degree than the Planck 
%team?

@@

% could compare temperatures from correlation slope analysis versus here, but 
% masks are different, so i guess it's just not worth it

@@

% mention that this is especially true because Planck gives such a huge
% increase in temperature resolution relative to SFD (and possibly reference
% temperature comparison figure as well)

@@

% make argument that, in models with f1 and the beta's constant, the choice
% of frequency at which tau is evaluated does not matter

@@

% MENTION THAT THE TAU TEMPLATE IS INDEPENDENT OF FREQUENCY FOR 
% TWO-COMPONENT MODELS WITH SPATIALLY CONSTANT BETAS, Q1/Q2 AND F1

@@

% could add a sentence praising the two-component reddening prediction for
% only having issues at the ~<10 mmag level for EBV_SFD ~< 0.4 mag, since
% it could be a lot worse ...

@@

However, we found that changing the zero level offsets of the input maps 
had a significant effect on the strength of the anticorrelation between median 
reddening residual and $T_2$. In particular, we experimented with perturbing 
the zero level offset of \PLANCK~857 GHz while correspondingly changing the 
zero levels of the remaining \PLANCK~maps based on the prescription of 
$\S$\ref{sec:relzero}. We also experimented with changing the zero level of SFD
\verb|i100|, independent of the other zero levels. Unfortunately, completely
flattening the reddening residual dependence on $T_2$ required unreasonably 
large zero level modifications. For example, flattening the $T_2$ residual 
required adding $\gtrsim$0.6 MJy/sr to the \verb|i100| map. Such an offset is 
implausible, being an order of magnitude larger than the nominal \verb|i100| 
zero level uncertainty quoted by SFD, and comparable to the entire 3000 GHz CIB
monopole signal. Furthermore, we note that even these large zero level 
modifications had virtually no effect in eliminating the reddening residual 
`bend' versus $E(B-V)_{SFD}$. Thus, changing the zero level offsets showed 
hints of promise in rectifying the reddening residual temperature dependence, 
but could not by itself completely resolve the systematic trends in reddening 
residuals.
% last few words here are kind of bad, need better terminology

@@

%%\section{Systematics}
%%
%%\subsection{CIBA}
%%Perhaps the dominant limitation of our fits is the presence of cosmic
%%infrared background anisotropies (CIBA) bleeding through into our maps of 
%%Galactic dust temperature and optical depth in high-latitude regions 
%%\citep{ciba}.
%%\subsubsection{Simulations}

%%\subsubsection{Optimizing Relative Weights}

%%\subsection{Zero Point}
%%Simulate the effect of getting zero points wrong to quantiy this
%%systematic a bit.

%%\subsection{$T_{2}$ Prior}
%%Use FIRAS data to justify $T_2$ prior. Could I use the DIRBE 100/240 micron
%%ratio to get a $1.3^{\circ}$ resolution prior on $T_2$?

%%\subsection{Other Limitations}
%%\IRAS~is missing in some parts of sky which will therefore not truly have 
%%6.1' resolution.
%%any other miscellaneous issues?

@@

% do i want to caution people not to extrapolate our 2comp model to higher
% frequencies, similar to what planck team did in their sec. 7.2.1 ??

% mention imprint of Planck team's half-degree on low-freq extrapolation ?

% is this the section where, potentially, plots of the residuals might go?

%%%It is true that we have smoothed the input intensity maps for this study 
%%%to $6.1'$ FWHM, which is nominally lower resolution than the $5'$ 
%%%\cite{planckdust} maps. However, as is apparent from figure 
%%%\ref{fig:comparison}, the fact that \cite{planckdust} modeled $\beta$ at 
%%%$0.5^{\circ}$ resolution has effectively blurred their dust temperature map.

@@

% if we're not going to be claiming to be superseding SFD in terms of 
% reddening, then do we even want to include anything extinction-related
% (data files and/or code) in this data release ?

@@

% in above paragraph, say which single-MBB beta(s) we find to be favored ?

@@

% sentence criticizing the half-degree beta single MBB model of Planck team
% in terms of the imprint that leaves on low-frequency emission 
% predictions ?

@@

% CIBA
The CIBA remains a major imperfection that still requires further 
investigation. The CIB anisotropies are very evident in low-dust regions of our
maps of optical depth and predicted dust emission. As described in 
$\S$\ref{sec:mcmc}, we have propagated the CIBA RMS amplitudes and 
inter-frequency covariances into our uncertainty estimates through the 
likelihood function in our MCMC procedure. However, this treatment falls far 
short of actually removing the spatial imprint of the CIBA on our derived 
parameters. The CIB anisotropies are more prominent in our optical depth map 
relative to that of SFD because of the lower-frequency \PLANCK~maps we rely 
upon to achieve a high-resolution temperature correction.

@@

% could add a sentence noting that similar zodi corrections are evident
% in all existing, comparable temperature map data products : Planck 2013a,
% SFD98 and FDS99

% audit for word ``zodi'' ... is this an acceptable jargon term ?
% could use IPD as an abbreviation at certain points, not sure if i've 
% defined that acronym already somewhere else in the text?

@@

% perhaps zodi will just magically be fixed in the forthcoming Planck
%   ---> do i want a figure illustrating the CIBA problem at high latitude ?
%  ---> zodi
%  ---> more testing needed, in any event, to prove Planck based reddening
%       estimates outperform SFD

% should mention that one definitive improvement of this work relative
% to FDS99 is that there are uncertainty/covariances provided for
% the derived parameters !!!!

% should there be a future work section or can the suggestions/plans of
% future work just be a part of ``SFD replacement'' subsection

% mention here the idea of getting even higher resolution by folding
% in WISE ?

>>>>

@@

@@

@@

@@

\section{Data}
\label{sec:data}
All \PLANCK~data products utilized throughout this work are drawn from the 
\PLANCK~2013 release \citep{planck2013}. Specifically, we have made use 
of all six of the zodiacal light corrected HFI intensity maps
\citep[\texttt{R1.10\_nominal\_ZodiCorrected},][]{planckzodi}. Our 
full-resolution (6.1$'$ FWHM) SED fits neglect the two lowest HFI frequencies, 
100 and 143 GHz, as these have FWHM of 9.66$'$ and 7.27$'$ respectively.

@@

The following subsections detail the processing steps we have applied to 
isolate Galactic dust emission in the \PLANCK~maps in preparation for SED
fitting.

@@

\subsection{Compact Sources}
\label{sec:ptsrc}
After subtracting the SMICA CMB model, we interpolate over compact sources, 
including both point sources and resolved galaxies. Removing compact sources at
this stage is important as it prevents contamination of compact-source-free 
pixels in our downstream analyses which require smoothing of the \PLANCK~maps. 
SFD carefully removed point sources and galaxies from the \verb|i100| map 
everywhere outside of $|b|$$<$$5^{\circ}$. We do not perform any further 
modifications of the \verb|i100| map to account for compact sources. To mask 
compact sources in the \PLANCK~217-857 GHz maps, we use the SFD compact source 
mask. At 100, 143 GHz we use the compact source masks provided by the 
\PLANCK~collaboration in the file \verb|HFI_Mask_PointSrc_2048_R1.10.fits|. 
Given our pixelization (see $\S$\ref{sec:pix}), 1.56\% of pixels are masked 
at 217-857 GHz (1.05\%, 1.02\% at 100, 143 GHz).

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

The color correction of Equation \ref{equ:bpcorr} therefore allows us to 
predict $\tilde{I}_{\nu_i}$ by computing monochromatic flux densities at the 
central frequency $\nu_{i,c}$ and then multiplying by factors 
$b_{\nu_i}(T, \beta)$. In practice, we interpolated the color corrections off 
of a set of precomputed, one-dimensional lookup tables each listing 
$b_{\nu_i}(T, \beta)$ for a single $\beta$ value as a function of $T$. We thus
avoided the need to interpolate in both $\beta$ and $T$ by computing only a 
small set of one dimensional correction factors for the particular set of 
$\beta$ values of interest (e.g. $\beta$=1.67, 2.7, 1.63, 2.82 ..., see Table 
\ref{tab:global}). This color correction approach makes the MCMC sampling 
described in $\S$\ref{sec:mcmc} much more computationally efficient by 
circumventing the need to perform the integral in the numerator of Equation 
\ref{equ:bpcorr} on-the-fly for each proposed dust temperature. We have chosen 
to compute the color corrections on a per-MBB basis because this approach is 
very versatile; all possible two-component (and single-MBB) models are linear 
combinations of MBBs, so we can apply all of our color correction machinery 
even when we allow parameters other than temperature (e.g. $f_1$) to vary and 
thereby modify the dust spectrum shape.

@@

\section{Global Model Parameters}
\label{sec:global}

@@

At 1250, 2141 and 3000 GHz, we use the SFD-reprocessed DIRBE maps. For each
DIRBE band, we determine $s_{i, 857}$ by performing a linear fit to DIRBE as a 
function of \PLANCK~857 GHz, after both have been zodiacal light subtracted and
smoothed to $1^{\circ}$ FWHM, also restricting to the low HI mask of 
$\S$\ref{sec:relzero} (see Figure \ref{fig:dirbe_slopes}).

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

However, we found that changing the zero level offsets of the input maps 
had a significant effect on the strength of the anticorrelation between median 
reddening residual and $T_2$. In particular, we experimented with perturbing 
the zero level offset of \PLANCK~857 GHz while correspondingly changing the 
zero levels of the remaining \PLANCK~maps based on the prescription of 
$\S$\ref{sec:relzero}. We also experimented with changing the zero level of SFD
\verb|i100|, independent of the other zero levels. Unfortunately, completely
flattening the reddening residual dependence on $T_2$ required unreasonably 
large zero level modifications. For example, flattening the $T_2$ residual 
required adding $\gtrsim$0.6 MJy/sr to the \verb|i100| map. Such an offset is 
implausible, being an order of magnitude larger than the nominal \verb|i100| 
zero level uncertainty quoted by SFD, and comparable to the entire 3000 GHz CIB
monopole signal. Furthermore, we note that even these large zero level 
modifications had virtually no effect in eliminating the reddening residual 
`bend' versus $E(B-V)_{SFD}$. Thus, changing the zero level offsets showed 
hints of promise in rectifying the reddening residual temperature dependence, 
but could not by itself completely resolve the systematic trends in reddening 
residuals.

@@

@@

@@

@@

@@

@@

The CIBA remains a major imperfection that still requires further 
investigation. The CIB anisotropies are very evident in low-dust regions of our
maps of optical depth and predicted dust emission. As described in 
$\S$\ref{sec:mcmc}, we have propagated the CIBA RMS amplitudes and 
inter-frequency covariances into our uncertainty estimates through the 
likelihood function in our MCMC procedure. However, this treatment falls far 
short of actually removing the spatial imprint of the CIBA on our derived 
parameters. The CIB anisotropies are more prominent in our optical depth map 
relative to that of SFD because of the lower-frequency \PLANCK~maps we rely 
upon to achieve a high-resolution temperature correction.

@@

@@
